Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance and associated polyandry

### 2017-03-18 ms_pre_post.Rmd
New round -- making clean edits with no comments for now. Change title and abstract

---
title: Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated
  polyandry
author: ''
date: "December 15, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding depression is widely hypothesised to drive adaptive evolution of pre-copulatory and post-copulatory mechanisms of inbreeding avoidance, which in turn are hypothesised to affect evolution of polyandry (i.e., female multiple mating). However, surprisingly little theory or modelling critically examines selection for pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, given evolutionary constraints and direct costs, or examines how evolution of inbreeding avoidance strategies might feed back to affect evolution of polyandry. Selection for post-copulatory but not pre-copulatory inbreeding avoidance requires polyandry, while interaction between pre-copulatory and post-copulatory inbreeding avoidance might cause functional redundancy (i.e., 'degeneracy') potentially generating complex evolutionary dynamics among inbreeding strategies and polyandry. We use individual-based modelling to quantify evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given strong inbreeding depression and different evolutionary constraints and direct costs. We find that evolution of post-copulatory inbreeding avoidance causes selection for increased polyandry, and that selection for a costly inbreeding avoidance strategy was weak or negligible given a non-costly alternative strategy. Further, fixed pre-copulatory inbreeding avoidance generally precluded evolution of polyandry and therefore post-copulatory inbreeding avoidance, but fixed post-copulatory inbreeding avoidance did not preclude pre-copulatory inbreeding avoidance from evolving. We thereby provide a general framework for understanding joint evolution of pre-copulatory and/or post-copulatory inbreeding avoidance, and conclude that evolution of inbreeding avoidance phenotypes and associated polyandry are predicted to be strongly affected by evolutionary feedbacks and degeneracy, and by direct costs on phenotypes affecting reproductive strategy.

>>>>

---
title: Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance and associated polyandry
author: ''
date: "December 15, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding depression is widely hypothesised to drive adaptive evolution of pre-copulatory and post-copulatory mechanisms of inbreeding avoidance, which in turn are hypothesised to affect evolution of polyandry (i.e., female multiple mating). However, surprisingly little theory or modelling critically examines selection for pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, given evolutionary constraints and direct costs, or examines how evolution of inbreeding avoidance strategies might feed back to affect evolution of polyandry. Selection for post-copulatory inbreeding avoidance but not for pre-copulatory inbreeding avoidance, requires polyandry, while interactions between pre-copulatory and post-copulatory inbreeding avoidance might cause functional redundancy (i.e., 'degeneracy') potentially generating complex evolutionary dynamics among inbreeding strategies and polyandry. We use individual-based modelling to quantify evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given strong inbreeding depression and different evolutionary constraints and direct costs. We find that evolution of post-copulatory inbreeding avoidance causes selection for increased polyandry, and that selection for a costly inbreeding avoidance strategy was weak or negligible given a cost free alternative strategy. Further, fixed pre-copulatory inbreeding avoidance generally precluded evolution of polyandry and therefore post-copulatory inbreeding avoidance, but fixed post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance. We thereby provide a general framework for understanding joint evolution of pre-copulatory and/or post-copulatory inbreeding avoidance, and conclude that evolution of inbreeding avoidance phenotypes and associated polyandry are predicted to be strongly affected by evolutionary feedbacks and degeneracy, and by direct costs on phenotypes affecting reproductive strategy.
